---
title: "assignement"
author: "Michelle Baumgarnter"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(epiwraps)
})
ah <- AnnotationHub()
ensdb <- ah[["AH89211"]]
```

## Download the data from encode for H3K4me3, H3K27me3 and smooth muscle cell peaks H3K27me3 and H3K4me3
```{r, eval = FALSE}
dir.create("peaks")
dir.create("tracks")
options(timeout =1800)

# H3K27me3
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz", "peaks/H3K27me3.bed.gz")
download.file("https://www.encodeproject.org/files/ENCFF160FEV/@@download/ENCFF160FEV.bigWig", "tracks/H3K27me3.bw")

# H3K4me3
download.file("https://www.encodeproject.org/files/ENCFF247GVM/@@download/ENCFF247GVM.bed.gz", "peaks/H3K4me3.bed.gz")
download.file("https://www.encodeproject.org/files/ENCFF611GSQ/@@download/ENCFF611GSQ.bigWig", "tracks/H3K4me3.bw")

# H3K27me3 smooth muscle cell peak
download.file("https://www.encodeproject.org/files/ENCFF369AEO/@@download/ENCFF369AEO.bed.gz", "peaks/smoothH3K27me3.bed.gz")

# H3K4me3 smooth muscle cell peak
download.file("https://www.encodeproject.org/files/ENCFF965WTD/@@download/ENCFF965WTD.bed.gz", "peaks/smoothH3K4me3.bed.gz")
```


# Reading in all the data from above
```{r}
k4me3 <- import("peaks/H3K4me3.bed.gz", format="NarrowPeak")
k27me3 <- import("peaks/H3K27me3.bed.gz", format="NarrowPeak")
smoothk27me3 <- import("peaks/smoothH3K27me3.bed.gz", format = "NarrowPeak")
smoothk4me3 <- import("peaks/smoothH3K4me3.bed.gz", format = "NarrowPeak")
```

# Finding bivalent domains between k4me3 and k27me3
```{r}
sum(overlapsAny(k4me3, k27me3))
```
We find 3219 bivalent domains between the two. 

```{r}
#Shows me the overlapping regions between k4m3 and k27me3
listofRegions <- list(k4me3=k4me3, k27me3=k27me3)
regionUpset(listofRegions)
```

# How many overlaps are there between k27m3 and smoothk27m3?
```{r}
sum(overlapsAny(k27me3, smoothk27me3))
```
We find 232 overlaps. 

# How many overlaps are there between k4me3 and smoothk4me3?
```{r}
sum(overlapsAny(k4me3, smoothk4me3))
```
We find 1542 overlaps

# Finding bivalent domains in the smooth muscle cells overlapping either mark or their combination. 
```{r}
smoothListofRegions <- list(k4me3=k4me3, k27me3=k27me3, smoothk27me3=smoothk27me3, smoothk4me3=smoothk4me3)

regionOverlaps(smoothListofRegions)
regionUpset(smoothListofRegions)
```

```{r}
sessionInfo()
```
 Time spent on this assignement ca. 2.5 hours