---
title: "02_Assignement"
author: "Michelle Baumgarnter"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Using AnnotationHub, find and download the following annotations data:
- The mouse (Mus Musculus) EnsDb object, version 102, genome build GRCm38
- The mouse genome sequence ( dna_sm ) in TwoBit/2bit format for GRCm38
- The drosophila melanogaster genome sequence ( dna_sm ) in TwoBit/2bit format for BDGP6

###Using the mouse EnsDb object, find the following:
- How many different ensembl gene IDs and gene symbols are there for protein-coding genes?
- Plot the distribution (histogram) of how many exons protein-coding genes
- Plot the distribution of the (spliced) length of protein-coding transcripts

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(ensembldb)
  library(GenomicRanges)
  library(epiwraps)
})
```

##Fetch gene annotation

```{r}
ah <- AnnotationHub()
#display(ah)
#Browse the AnnotationHub to find the Mus Musculus GRCm38 version 102...  
Mus102 <- ah[["AH89211"]]
Mus102
#...mouse genome sequence (dna_sm) in 2bit format for GRCm38...
Musdna_sm <- ah[["AH88477"]]
Musdna_sm
#...and Drosophila melanogaster genome sequence (dna_sm) in two bit format for BDGP6. 
Drodna_sm <- ah[["AH68180"]]
Drodna_sm
```

##How many different ensembl gene IDs and gene symbols are there for protein-coding genes?
```{r}
#Shows me all the possible data types which can be retrieved from Mus102
columns(Mus102)
listTables(Mus102)
listGenebiotypes(Mus102)
MusGen <- genes(Mus102, filter = GeneBiotypeFilter("protein_coding"), columns=c("gene_id", "symbol"))
#Shows me all the protein coding gene_id and symbols
MusGen
#Number of different gene_id and symbols
length(MusGen$gene_id)
length(MusGen$symbol)
```



##Plot the distribution (histogram) of how many exons protein-coding genes
```{r}
#Exons are sorted by chromosome, strand, start and end values by using by = gene
exprocod <- exonsBy(Mus102, by = "gene", filter=GeneBiotypeFilter("protein_coding"))
numberexprocod <- lengths(exprocod)
numberexprocod
hist(numberexprocod)

```

##Plot the distribution of the (spliced) length of protein-coding transcripts
```{r}
lengthproco <- sum(width(exprocod))

hist(lengthproco)
```
##Time spent on this assignement ca. 4.5 hours

```{r}
sessionInfo()
```