---
title: "assignement"
author: "Michelle Baumgarnter"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(memes)
  library(Biostrings)
  library(AnnotationHub)
  library(MotifDb)
  library(universalmotif)
  library(PWMEnrich)
  library(ggplot2)
})
```
```{r}
#make sure to have the updated version
BiocManager::install(c("MotifDb", "universalmotif", "TFBSTools", "memes", "PWMEnrich"))
```

Choose a transcription factor, e.g. GATA1, MYOD1, CEBPB, REST (or any of your choice that has a motif and available ChIPseq data)
Download the (e.g. Mouse) peaks for that factor (whatever cell type)
```{r}
download.file("https://www.encodeproject.org/files/ENCFF366HNL/@@download/ENCFF366HNL.bed.gz", "mESC_GATA1_ENCFF366HNL.bed.gz")
peaks <- rtracklayer::import("mESC_GATA1_ENCFF366HNL.bed.gz", format = "NarrowPeak")
seqlevelsStyle(peaks) <- "Ensembl"
sum(table(peaks))
```

Identify the instances of the factor’s motif
```{r}
#gives us slighly shorter sequence to make search for motif easier
peak_centers <- resize(peaks, fix="center", width=100)
# we get the genome sequence:
ah <- AnnotationHub()
genome <- ah[["AH68356"]]
# we get the sequences corresponding to the peak centers:
peak_seqs <- memes::get_sequence(peak_centers, genome)
peak_seqs
```

```{r}
#' findMotifInstances
#'
#' A wrapper around TFBSTools for scanning motif occurence, handling the 
#' coordinate conversion as `memes` does.
#' 
#' @param seqs A set of sequences, e.g. `DNAStringSet`, optionally with 
#'   coordinate as names to enable conversion.
#' @param motif A motif, in any format recognized by `universalmotif`
#' @param keepMatchedSeq Logical; whether to keep the matched sequence.
#' @param ... Passed to `TFBSTools::searchSeq`; can for instance be used to set
#'   the number of threads to use, e.g. with `mc.cores=2`
#'
#' @return A `GRanges` object
findMotifInstances <- function(seqs, motif, keepMatchedSeq=FALSE, ...){
  motif <- universalmotif::convert_motifs(motif, "TFBSTools-PWMatrix")
  x <- suppressWarnings(TFBSTools::searchSeq(motif, subject=seqs))
  peaks <- strsplit(gsub("-",":",names(seqs)), ":")
  if(all(lengths(peaks)==3)){ # convert relative coordinates to absolute
    chrs <- sapply(peaks,FUN=function(x) x[1])
    offsets <- sapply(peaks,FUN=function(x) as.integer(x[2]))-1L
    i <- rep(seq_along(x),lengths(x))
    return(GRanges(chrs[i], 
                   IRanges(offsets[i]+as.integer(unlist(lapply(x, start))),
                           offsets[i]+as.integer(unlist(lapply(x, end)))),
                   strand=unlist(lapply(x, strand)),
                   score=as.numeric(unlist(lapply(x, FUN=function(x) score(x))))
                   ))
  }
  x <- as(x, "GRanges")
  score(x) <- x$absScore
  keepFields <- c("score","relScore")
  if(keepMatchedSeq) keepFields <- c(keepFields, "siteSeqs")
  mcols(x) <- mcols(x)[,intersect(colnames(mcols(x)), keepFields)]
  x
}
```
```{r}
# we search for "GATA1" in the motif database, so we check what are my available motifs and then I can choose one, if Jaspar has one I choose this one because it has high quality if not this available I will choose HOCOMOCO, quality can be seen on Letter, number at end
motifs <- query(MotifDb, "GATA1")
# there are several matching motifs:
names(motifs)
# we select one:
motif <- motifs[["Mmusculus-jaspar2018-Gata1-MA0035.1"]]
# we visualize it:
view_motifs(motif)
```
```{r}
motif_instances <- findMotifInstances(peak_seqs, motif, mc.cores=2) # running with 2 threads
```
```{r}
mmusculus <- import(genome)["1"]
motif_instances_genome <- findMotifInstances(mmusculus, motif, mc.cores=2)
```

```{r}
sum(overlapsAny(motif_instances, peaks))
percent <- (sum(overlapsAny(peaks, motif_instances)))/length(peaks)*100
```
```{r}
motifs_with_peaks <- overlapsAny(motif_instances_genome, peaks)
sum(motif_with_peaks)
percent1 <- sum(motifs_with_peaks/length(motif_instances_genome))*100
```

Answer the following questions:
Of all the peaks, what proportion contains a motif for the factor? 
- Of the `r length(peaks)`, `r sum(overlapsAny(peaks, motif_instances))` (`r percent`%) contain a motif. 

Of all instances of that motif in the genome, what proportion is bound by the factor (i.e. has a peak)? 
Of the `r length(motif_instances_genome)`motif instances, `r sum(motif_with_peaks)` (`r percent1`%) overlap a peak. 

```{r}
sessionInfo()
```

Time spent on this assignement ca. 3h.